---
title: "79546 - Timeseries Emphirical Paper"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

Clear all variables in work space and install packages

```{r message = FALSE,warning = FALSE}
rm(list=ls())

requiredPackages = c('quantmod','TTR','tseries')
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p) #install package if it does not exist
  library(p,character.only = TRUE)
}

#suppress `getSymbols` message
options("getSymbols.warning4.0"=FALSE)

```

## Load the forecasting packages

```{r warning=FALSE}
library(fpp2)
library(TTR)
library(quantmod)
library(ggplot2)
library(tseries)

```

## Get to Know Time-Series Data
### Load Dataset

Data is collected from Yahoo Finance using the Quantitative Financial Modeling Framework (Quantmod). Data obtained in eXtensible-Time-Series format is being used for data exploration.

```{r  }
#Download data from yahoo finance
df_tsm <- getSymbols('TSM',src='yahoo',auto.assign=FALSE,from="2011-01-01")

#Check the contents of the data
class(df_tsm)

#List the number of rows in the data
nrow(df_tsm)

#Print the last 6 rows of the data
tail(df_tsm)


```

### Stock price visualization
This shows the patterns of the data.

```{r}
tsm_title = "Taiwan Semiconductor Manufacturing Company Limited Stock Price (TSM) (2011-2022)"


chartSeries(df_tsm , name="TSM price 2011-2022")

```
TSM share price has grown upwards from 2011 to 2022 

### Time plot of the data

```{r }
#returns the closing price
tsm_close = Cl(to.monthly(df_tsm))

#decompose the data
dc <- decompose(as.ts(tsm_close, start=c(2011,1)))
plot(dc)

#Display seasonal stock data
dc$seasonal

```
The output shows for plots of TSM closing price which are:  
- **Observed**: Original plot of the data.   
- **Trend** : There is an upward trend that is significant from 2018.   
- **Seasonal** There is repetitive seasonal fluctuation of data.The closing price reached the highest in January and the lowest in June.This shows that a good time to sell is beginning of the year and the risht time to buy is mid year  
- **Random** irregular or random fluctuation not captured by the trend and seasonal.  


```{r}
#plot(df_tsm$TSM.Close,main = tsm_title)
chart_Series(tsm_close,name=tsm_title)

```
From the figure above TSM stock price has a `strong positive trend`.This shows that it is `non-stationary`

### Stationarity Test on the stock price

```{r}
adf.test(tsm_close)
```
The p-value is not less than 0.05 hence we fail to reject null hypothesis.  
This means that the time series is non-stationary.

### ACF and PACF plots of the time-series data

```{r}

par(mfrow=c(1,2))
acf(tsm_close)
pacf(tsm_close)

```
from the ACF and PACF plot,AR model would be ideal for this stock price  
The trend can be removed by `differensing ` the data to removes the trend
```{r}
par(mfrow=c(2,1))

dy = diff(tsm_close,lag = 1)
chart_Series(dy,name="Time Plot of difference data")

wld = diff(log(tsm_close))
chart_Series(wld,name="Time Plot of log difference data")


```

### Stationarity test of Differencing log time-series

```{r}
adf.test(wld[!is.na(wld)])

```
The p-value is less than 0.05 hence we accept null hypothesis.  
This means that the difference time series is stationary.

## Find the best fit arima model
`auto.arima` finds the best fit ARIMA model the forecasting
```{r}
# differencing is set to 1 d=1
# if TRACE = TRUE prints out all models that have been tried
fit_arima = auto.arima(tsm_close,d=1,stepwise = FALSE,approximation = FALSE,trace = FALSE)
print(summary(fit_arima))
checkresiduals(fit_arima,plot=TRUE)
```
### Genarate a 24-month forecast using best fit ARIMA model
```{r}

fcst =  forecast(fit_arima, h=24,level=c(95))
autoplot(fcst)

```